---
title: "dlyr102data"
format: html
editor: visual
author: Disem Sula
---

```{r}
#| echo: false
#| message: false
here::i_am("dplyr-102.Rproj")
library(here)
library(vroom)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
```

### Financial data analasys

```{r}
#| echo: false
#| message: false
clients <- vroom("data/client.csv")
accounts <- vroom("data/account.csv")
disp <- vroom("data/disp.csv")
```

## data tables

A collection of joining functions of the form `*_join`

### Clear data

```{r}
da <- inner_join(accounts, disp)
```

```{r}
cda <- inner_join(clients, da, by=join_by(client_id), suffix = c("_home", "_bank"))
```

```{r}
home_diff_bank <- cda |> filter(district_id_home != district_id_bank) |> nrow()
```

We observe that `r home_diff_bank` clients have a bank to another

```{r}
cda |>
  group_by(account_id) |> 
  distinct(district_id_home) |> 
  summarise(N=n()) |> 
  filter(N>1)
```

### Gender gap

```{r}
clients |> count(gender)
```

```{r}
clients |> 
  ggplot(aes(x=gender)) + geom_bar() + coord_flip()
```

```{r}
cda |> 
  count(gender, type) |>
  group_by(gender) |>
  mutate(freq=n/sum(n)) |>
  select(-n) |> 
  ungroup() |>
  pivot_wider(values_from = freq, names_from = type)
```

```{r}
cda |>
  ggplot(aes(y=gender, fill=type)) + geom_bar(position = "fill")
```

```{r}
chisq.test(cda |> pull(gender),
           cda |> pull(type))
```

## Loan analysis
```{r}
#| message: false
loan <- vroom("data/loan.csv")
```

```{r}
#mutate starts mutating from the first column on.
#we have to do the modifications on the data when loading if it's a small dataset, or in a different chunk if it's big
loan <- loan |>
  mutate(status = factor(status, levels=c( "A", "B", "C", "D")), loan_id = as.integer(loan_id), account_id=as.integer(account_id))
```

Factor recording of `status` introduced `r loan |> filter (is.na(status)) |> nrow()` error(s)

```{r}
loan_counts <- ggplot(loan, aes(status)) +
  geom_bar()
loan_counts
```

```{r}
loan |> group_by(status) |>
  summarise(sum(amount))
```

```{r}
ggplot(loan, aes(status, amount)) +
  geom_boxplot()
```

```{r}
loan_amounts <- loan |> group_by(status) |>
  summarise("total amount" = sum(amount)) |>
  ggplot(aes(x=status, y=`total amount`)) +
  geom_col()
loan_amounts
```

```{r}
loan_counts + loan_amounts
```

```{r}
loan_counts / loan_amounts
```

```{r}
(loan_counts + coord_flip()) / (loan_amounts + coord_flip())
```

```{r}
(loan |> 
  filter(status %in% c("A", "B")) |>
  ggplot(aes(status)) + geom_bar()) +
(loan |> 
  filter(status %in% c("C", "D")) |>
  ggplot(aes(status)) + geom_bar())
```

```{r}
loan <- loan |>
  mutate(active=if_else(status %in% c("A", "B"), "Past", "Active"))
```

```{r}
ggplot(loan, aes(x=status)) +
  geom_bar() +
  facet_wrap(~active, scales = "free")
```

```{r}
ggplot(loan, aes(y=active, fill=status)) +
  geom_bar()
```

```{r}
loan <- loan |>
  mutate(default = status %in% c("B", "D"))
```

```{r}
ggplot(loan, aes(y=active, fill=status)) +
  geom_bar(position="fill")
```

## Combined analysis

```{r}
loan_cda <- loan |>
  inner_join(cda, by=join_by(account_id), suffix=c("_loan", "_account"))
```

```{r}
loan_cda |>
  group_by(loan_id) |>
  summarise(nb_clients = n())
```

```{r}
loan_cda |>
  group_by(loan_id) |>
  mutate(nb_clients = n()) |>
  slice_head() |>
  select(-client_id, -gender, -birth_date) |>
  ungroup() |>
  ggplot(aes(y=nb_clients, fill=default)) +
  geom_bar()
```

We see that when we have two persons owning the account we have no dafaults.

```{r}
client_no_loans <- cda |> 
  anti_join(loan, by=join_by(account_id))
client_with_loans <- cda |> 
  semi_join(loan, by=join_by(account_id))
```
We find clients who do not have a loan, anti_join picks the values in x that do not have a match in y, with semijoin I take the ones that have a match, with inner all together

```{r}
(client_no_loans |>
  group_by(account_id) |>
  summarise(nb_clients=n()) |>
  ggplot(aes(y=nb_clients)) +
  geom_bar() + ggtitle("Without loan")) /
(client_with_loans |>
  group_by(account_id) |>
  summarise(nb_clients=n()) |>
  ggplot(aes(y=nb_clients)) +
  geom_bar() + ggtitle("With loan"))
```

```{r}
merged_clients <-
  bind_rows(
    with = client_with_loans,
    without = client_no_loans,
    .id = "loan")
```

```{r}
merged_clients |>
  group_by(account_id) |>
  mutate(nb_clients =n()) |>
  slice_head() |>
  ungroup() |>
  ggplot(aes(y=nb_clients, fill=loan)) +
  geom_bar()
```
There's difference into counts so we need fill:
```{r}
merged_clients |>
  group_by(account_id) |>
  mutate(nb_clients =n()) |>
  slice_head() |>
  ungroup() |>
  ggplot(aes(y=nb_clients, fill=loan)) +
  geom_bar(position="fill")
```

```{r}
accounts_loan <- merged_clients |>
  group_by(account_id) |>
  mutate(nb_clients =n()) |>
  slice_head() |>
  ungroup()
chisq.test(accounts_loan |> pull(loan),
           accounts_loan |> pull(nb_clients))
```
### Alternative solution

```{r}
persons_per_account <- cda |>
  count(account_id)
loan |> inner_join(persons_per_account, by=join_by(account_id))
```

```{r}
loan_cda |>
  group_by(default, loan_id) |>
  summarise(n=n(), .groups="drop")
```

```{r}
cda |> 
  left_join(loan|>select(account_id, default), by=join_by(account_id)) |>
  mutate(loan = if_else(is.na(default), "Without", "With"))
#left is "cda" and join it with loan on the "right"
```


